---
title: "KV_Assignment4_fml"
author: "Keerthana Vonteddu"
date: "2024-03-17"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Summary
1. The data has been standardized using range and Z-score scaling.
2. Five clusters with five centroids are the optimal outcome when k-means clustering is carried out using the silhouette technique.
3. The best k = 5 number we came up with in this instance is roughly 0.3, which is good. Higher values indicate higher quality. The silhouette ranges from -1 to +1.
4. The cluster names are categorized using the properties of the data set and the cluster summary:
NetProfitMargin was higher in Cluster 1 and "High Market Cap & High ROE"
Hold Cluster refers to Cluster 2.
The third cluster is "Low Market Cap & Less ROA".
The fourth cluster is "Small Net Profit Margin & High PE Ratio."
The fifth cluster is "Moderate Buy & Moderate Sell".
#Problem Statement:
An equities analyst is researching the pharmaceutical sector and needs your assistance in examining and deciphering the financial information that her company has gathered. Her primary goal is to use certain fundamental financial metrics to comprehend the pharmaceutical industry's structure. The Pharmaceuticals.csv file contains financial information collected on 21 pharmaceutical companies. Get the Pharmaceuticals.csv file.
The following variables are noted for every firm.

1. Market capitalization expressed in USD per billion. 2. Beta. The ratio of price to earnings. 4. Equity return. 5. Asset return. 6.Asset turnover. 7. Make use of leverage. 8. Projected growth in sales. 9. The net profit margin. 10. Median advice (for all major brokerages). 11. The location of the corporate office. 12. The stock exchange where the company is listed.

#Packages Required
```{r}
#install.packages("flexclust")
library(caret)
library(ISLR)
library(dplyr)
library(factoextra)
library(flexclust)
library(tidyverse)
library(readr)
```

#keeping in mind that KV_A4_fml is our dataframe and loading information about medications
```{r}
KV_A4_fml <- read.csv("/Users/keerthanavonteddu/Desktop/fml/Pharmaceuticals.csv")
summary(KV_A4_fml)
head(KV_A4_fml)
```

##Q1.Use only the numerical variables (1 to 9) to cluster the 21 firms. Justify the various choices made in conducting the cluster analysis, such as weights for different variables, the specific clustering algorithm(s) used, the number of clusters formed, and so on
```{r}
data_scaled.df <- scale(KV_A4_fml[,3:11])
head(data_scaled.df)
summary(data_scaled.df)
```

#Distance Measurement in Data Analysis for Scaled Medicine
```{r}
distance.df <- get_dist(data_scaled.df)
fviz_dist(distance.df)
```

#Scale the data and arrange it using K-means clustering.
```{r}
kmeans._1.df <- kmeans(data_scaled.df,centers = 5,nstart = 25)
kmeans._1.df

fviz_cluster(kmeans._1.df,data = data_scaled.df)  #graphical Representation of the Cluster
```

Overview: Information on nine attributes pertaining to each of the twenty-one pharmaceutical businesses can be acquired by applying the summary algorithm on the k-means data. There are five different clusters in total, each having five centroid points: 2, 3, 4, 8, and 9.

##Q2.Interpret the clusters with respect to the numerical variables used in forming the clusters

```{r}
#graphing the total number of clusters versus the sum of the values
fviz_nbclust(data_scaled.df,kmeans,method = "wss")
```

This figure makes it evident that, rather than taking the form of an elbow, the graph is significantly flattening at k=4 and k=6.

#By graphing the number of clusters against the typical silence width, you can determine the ideal number of clusters.
```{r}
fviz_nbclust(data_scaled.df,kmeans,method = "silhouette")
```

We can quickly determine that five clusters is the best and most effective option for this k-means clustering by using the Silhoutte graph.

##Q3.Is there a pattern in the clusters with respect to the numerical variables (10 to 12)? (those not used in forming the clusters).

#utilizing the data shown in the clusters to get the mean value.
```{r}
#executing the aggregate functionality of the pharmaceutical data.
aggregate(KV_A4_fml[3:11],by=list(cluster = kmeans._1.df$cluster),mean)
```

```{r}
#Use Cbind to merge the data frames.
shot.data <- cbind(KV_A4_fml,cluster = kmeans._1.df$cluster)
tibble(shot.data)
```

#recapitulating the detailed division by cluster
```{r}
by(shot.data,factor(shot.data$cluster),summary)
```
#Median the calculation
```{r}
Recm.table <- table(shot.data$cluster,shot.data$Median_Recommendation)
names(dimnames(Recm.table)) <- c("Cluster","Recommendation")
Recm.table <- addmargins(Recm.table)
Recm.table
```
There are 21 suggestions total: 4 holds, 7 moderate buys, 4 moderate sells, and 1 strong buy. Cluster 5 combines the purchase and sell recommendations that are in opposition to the other three. Clusters 1, 2, and 3 contain all of the information about mod purchases and holds.For Cluster 4, there is a moderate buy and a moderate sell recommendation.

#A cluster breakdown of the organization's headquarters based on the aggregated data.
```{r}
location_table <- table(shot.data$cluster,shot.data$Location)
names(dimnames(location_table)) <- c("Cluster","Location")
location_table <- addmargins(location_table)
location_table
```
In total, there are 21 companies: 13 are located in the United States, 3 in the United Kingdom, and 1 in each of Canada, France, Germany, Ireland, and Switzerland. Group 5 includes the United States, the United Kingdom, and Switzerland. Cluster 2 includes the US and Canada. Cluster 3 includes the US and Britain, whereas Cluster 4 includes France, Ireland, and the US.

#making a data frame with the combined data and building the exchange table to show a summary of the stock exchange values for each cluster.
```{r}
exc_table <- table(shot.data$cluster,shot.data$Exchange)
names(dimnames(exc_table)) <- c("Cluster","Exchange")
exc_table <- addmargins(exc_table)
exc_table
```
There are a total of 21 companies: 1 Amex, 1 Nasdaq, and 19 NYSE. The NYSE is the lone entity in Cluster 5. Cluster 2 comprises all three. The sole organization in clusters 1, 3, and 4 is NYSE.
##Q3. Provide an appropriate name for each cluster using any or all of the variables in the dataset.

Cluster 1: "High Market Cap - High ROE" greater Net-profit Margin is made up of all US companies listed on the NYSE that are either held or acquired.

Cluster 2 is known as the "Hold Cluster" since it only consists of companies that are listed on the New York Stock Exchange.

Cluster 3: On stock exchanges, this is referred to as "Low Market Cap & Less ROA" and is either buy or hold.

Cluster 4: "High PE ratio / Small Net Profit Margin" is the category this falls under.

Cluster 5: This cluster may be referred to as "Moderate Buy & Moderate Sell."


#Range the scaling
```{r}
data_range <- scale(KV_A4_fml[,3:11])
summary(data_range)
head(data_range)
```

#trying out or researching various techniques to see if they could also provide more effective clustering?
```{r}
fviz_nbclust(data_range,FUN = kmeans,method = "silhouette")
kmeans_2.df <- kcca(data_scaled.df,k=5,kccaFamily("kmeans"))
kmeans_2.df
clusters(kmeans_2.df)
```
#Using the predict() function.
```{r}
index_cluster.df <- predict(kmeans_2.df)
image(kmeans_2.df)
points(data_scaled.df,col = index_cluster.df,pch = 9,cex = 1.0)
```
Here, the kcca method is utilized to perform a kmeans cluster on k = 5, as opposed to the kmeans function from basic R. While the clustering has the same size as the underlying R approach, the assignment between the points is different. The clustering graph shows that there is more ambiguity in the categorization than we would like, especially between groups 1, 2, and 3.

#PLOTTING K-MEDIANS AND CLUSTERS.
```{r}
kmeans_2.df <- kcca(data_scaled.df,k = 5,kccaFamily("kmedians"))
kmeans_2.df
clusters(kmeans_2.df)
index_cluster.df <- predict(kmeans_2.df)
image(kmeans_2.df)
points(data_scaled.df,col = index_cluster.df,pch = 9,cex = 1.0)
```
The five clusters have the following sizes when the KCC is converted from Kmeans to Kmedian: 4, 5, 3, 3, and 6.However, the clustering is less clear. Even though it's unclear if a better cluster exists, we want to look over the new data to see if there are any other methods or tools we might utilize to improve the visual cluster.


















































