---
title: "FML assignment 2"
author: "Keerthana"
date: "2024-02-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***
#Summary
1.Data Preprocessing:

The code reads the dataset from a CSV file and removes the "ID" and "ZIP Code" columns.
It converts the "Education" column into a factor variable and creates dummy variables for it.

2.Data Partitioning:

The dataset is split into training (50%), validation (30%), and test (20%) sets using createDataPartition.

3.Normalization:

The numeric features in the dataset are normalized using z-score normalization (preProcess function).

4.KNN Model Training and Evaluation:

A KNN classification model is trained using the training data with k = 3.
The model's predictions are evaluated using confusion matrices on the training, validation, and test sets.
The confusion matrices provide insights into the model's performance in terms of true positives, true negatives, false positives, and false negatives.

5.Interpretation:

The confusion matrices provide a summary of the model's predictive performance.
Metrics such as accuracy, precision, recall, and F1-score can be derived from the confusion matrices to further evaluate the model.
Based on the confusion matrix for the test set, the model achieved a certain number of true positives, true negatives, false positives, and false negatives, indicating its performance in predicting loan acceptance correctly and incorrectly.
The same evaluation is done for the training and validation sets to ensure the model's generalization ability.
The code concludes with interpreting the results and summarizing the performance of the KNN model.

##Here I am Importing the Dataset
```{r}
bank.df= read.csv("/Users/keerthanavonteddu/Desktop/fml/UniversalBank.csv")
```
```{r}
summary(bank.df)
```

```{r}
library(caret)
library(class)
library(dplyr)
library(ISLR)
library(lattice)
```
##Now I am going to Remove ID ,  ZIP Code

```{r}
bank.df$ID <- NULL
bank.df$ZIP.Code <- NULL
bank.df$Education = as.factor(bank.df$Education)

```
##later I am creating a dummy dataset
```{r}
# Using model.matrix() to create dummy variables
dummyvari <- as.data.frame(model.matrix(~ bank.df$Education - 1))

```

```{r}
names(dummyvari) <- c("Education_1", "Education_2","Education_3")
```


##Now I am setting  education to NULL
```{r}
bank.df$Education <- NULL
```

```{r}
bank_Final <- cbind(bank.df, dummyvari)
```

##Here I am going to divide the dataset into train and test
```{r}
set.seed(1)
train.index <- createDataPartition(bank_Final$Personal.Loan, p= 0.6 , list=FALSE)
valid.index <- setdiff(row.names(bank_Final), train.index)
train.df <- bank_Final[train.index,]
valid.df <- bank_Final[valid.index,]

```
##Now we are generating the Test data
```{r}
new_customer <- data.frame(Age = 40,
                       Experience = 10,
                       Income = 84,
                       Family = 2,
                       CCAvg = 2,
                       Mortgage = 0,
                       Securities.Account = 0,
                       CD.Account = 0,
                       Online = 1,
                       CreditCard = 1,
                       Education_1 = 0, 
                       Education_2 = 1, 
                       Education_3 = 0)
```

## normalisation

```{r}
train.norm.df <- train.df[,-7]
valid.norm.df <- valid.df[,-7]
new_customer.norm <- new_customer
norm.values <- preProcess(train.df[, -7], method=c("center", "scale"))
train.norm.df <- predict(norm.values, train.df[, -7])
valid.norm.df <- predict(norm.values, valid.df[, -7])
new_customer.norm <- predict(norm.values, new_customer.norm)
```
```{r}
summary(train.norm.df)
```

##Here I am Performing Knn classification, using K=1
```{r}

outputrate <- class::knn(train = train.norm.df,test = new_customer.norm,
                       cl = train.df$Personal.Loan, k = 1)

print(outputrate)

```

##we are going to find best K

```{r}
accuracy.df <- data.frame(k = seq(1, 10, 1), accuracy = rep(0, 10))
```

```{r}
for(i in 1:10) {
  knn.prediction <- class::knn(train = train.norm.df,
                         test = valid.norm.df,
                         cl = train.df$Personal.Loan, k = i)
  accuracy.df[i, 2] <- confusionMatrix(knn.prediction,
                                       as.factor(valid.df$Personal.Loan))$overall[1]
}
which(accuracy.df[,2] == max(accuracy.df[,2]))
```

```{r}
accuracy.df
```

##choosing k = 3
```{r}
knn.prediction <- class::knn(train = train.norm.df,
                       test = valid.norm.df,
                       cl = train.df$Personal.Loan, k = 3)

confusionMatrix(knn.prediction, as.factor(valid.df$Personal.Loan), positive = "1")
```
##Now Here confusion matrix for the best k value =3

```{r}
newcustomer <- data.frame(Age = 40,
                            Experience = 10,
                            Income = 84,
                            Family = 2,
                            CCAvg = 2,
                            Mortgage = 0,
                            Securities.Account = 0,
                            CD.Account = 0,
                            Online = 1,
                            CreditCard = 1,
                            Education_1 = 0, 
                            Education_2 = 1, 
                            Education_3 = 0)

fitknn <-class::knn(train = train.norm.df,
                          test = newcustomer,
                          cl = train.df$Personal.Loan, k = 3)

fitknn
```

##knn model tells that new customer will accept loan
##reading the dataset

```{r}
bank.df= read.csv("/Users/keerthanavonteddu/Desktop/fml/UniversalBank.csv")
```

##packages

```{r}
library(ISLR)
library(caret)
library(class)
library(dplyr)
library(lattice)

```
##here i am removing id and zipcode variables from the dataset

```{r}
bank.df$ID <- NULL
bank.df$ZIP.Code <- NULL
bank.df$Education = as.factor(bank.df$Education)
```

##Creating a dummy dataframe

```{r}
# Using model.matrix() to create dummy variables
dummy_matrix <- model.matrix(~ bank.df$Education - 1)
dummymod <- as.data.frame(dummy_matrix)
 
```
##Renaming the data frame

```{r}
names(dummymod) <- c("Education_1", "Education_2","Education_3")
```



##deleting education variable

```{r}
bank.df$Education <- NULL 
```
##Main dataset

```{r}
bank_Final <- cbind(bank.df, dummymod)
```
#Partitioning  the dataset

```{r}
set.seed(1)
train.index <- createDataPartition(bank_Final$Personal.Loan, p= 0.5 , list=FALSE)
valid.index <- createDataPartition(bank_Final$Personal.Loan, p= 0.3 , list=FALSE)
test.index <- setdiff(row.names(bank_Final), union(train.index, valid.index))
                      
train.df <- bank_Final[train.index, ]
valid.df <- bank_Final[valid.index, ]
test.df <- bank_Final[test.index, ]

```
##Performing normalisation

```{r}

train.norm.df <- train.df[,-7]
valid.norm.df <- valid.df[,-7]
test.norm.df <- test.df[,-7]

norm.values <- preProcess(train.df[, -7], method=c("center", "scale"))
train.norm.df <- predict(norm.values, train.df[, -7])
valid.norm.df <- predict(norm.values, valid.df[, -7])
test.norm.df <- predict(norm.values, test.df[, -7])

```
##Performing Knn classification using K=3

```{r}
knn.test.pred <- class::knn(train = train.norm.df,
                       test = test.norm.df,
                       cl = train.df$Personal.Loan, k = 3)

knn.train.pred <- class::knn(train = train.norm.df,
                             test = train.norm.df,
                             cl = train.df$Personal.Loan, k = 3)

knn.valid.pred <- class::knn(train = train.norm.df,
                             test = valid.norm.df,
                             cl = train.df$Personal.Loan, k = 3)
```
##Confusion matrix for K=3

```{r}
confusionMatrix(knn.test.pred, as.factor(test.df$Personal.Loan), positive = "1")
confusionMatrix(knn.train.pred, as.factor(train.df$Personal.Loan), positive = "1")
confusionMatrix(knn.valid.pred, as.factor(valid.df$Personal.Loan), positive = "1")
```






